Bayesian statistics and swine flu

### 2009-08-12 docs/pandemic/bayesPandemic.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@318 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@354 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass{beamer}
\usepackage[nogin]{Sweave}
\title{Bayesian statistics and  swine flu}
\author{Patrick Brown, Dionne Gesink Law, Peter Neal} 
\usepackage{amsmath,subfigure}

@@

<<setup,fig=false>>=
options(SweaveHooks=list(fig=function() par(mar=c(2.5,2.5,0.1,0.1), mgp=c(1.5, 0.5, 0), cex=0.6)))
# source in all the R files
Rfiles=system("CMD /K dir /B ..\\..\\pkg\\pandemic\\R\\*.R", intern=T)
Rfiles = grep("\\.R$", Rfiles, value=T)
for(D in Rfiles) {
	source(paste("..\\..\\pkg\\pandemic\\R\\", D, sep=""))
#$
}
@

@@

@@

\begin{block}{Mild}
\begin{center}
 Infection $\stackrel{f_I(t)}{\Longrightarrow}$ Onset $\stackrel{f_{OM}(t;)}{\Longrightarrow}$ \\ Medical  $\stackrel{f_{MR}(t)}{\Longrightarrow}$ Recovery
\end{center}

@@

\begin{frame}
\frametitle{Distributions}
\begin{itemize}
	\item $f(t;\mu,\nu,\delta)$ is the distribution for time to an event.  
	\item \ldots a rounded, zero-inflated Weibull
\begin{align*}
Z\sim&\text{Bern}(\delta)\\
[Y|Z=1] =& 0 \text{w.p.} 1\\
pr(T=t|Z=0) =&  \text{pr}(\max(t-0.5,0) < y < t+0.5)\\
y \sim & \text{Weibull}(\mu/\Gamma(1 + 1/\nu), \nu)
\end{align*}
\item $\delta$ is the probability of transitioning immediately for an administrative (not biological) reason, 
\begin{itemize}
	\item i.e.\ arriving dead at the hospital.
\end{itemize}
\item Barring administrative zeros, $\mu = E(T|Z=0)$ is the average time to an event
\begin{itemize}
	\item $\mu_I$ is average time to infection
\end{itemize}
\item $\nu$ is a Weibull shape parameter
\begin{itemize}
	\item $\nu=1$ is constant hazards
	\item $\nu > 1$ results in more regular event times.
\end{itemize}

@@

\begin{frame}
\frametitle{Prior graphs}
<<priorGraph, fig=false,results=tex>>=
priors= pandemicPriors()
plotPrior(priors$InfOns, file="InfOns", tex=TRUE)
#$
@

@@

\begin{frame}
\frametitle{To do}
\begin{itemize}
	\item model the infection process
	\item simulate hospital loads
	\item covariates
	\item spatial infection process
\end{itemize}
\end{frame}

>>>>

\documentclass{beamer}
\usepackage[nogin]{Sweave}
\title{Bayesian statistics and  swine flu}
\author{Patrick Brown$^1$,  Peter Neal$^2$}
\institute{1:Cancer Care Ontario and Dalla Lana School of Public Health, University of Toronto; 2: University of Manchester} 
\usepackage{amsmath,subfigure}

@@

<<setup,fig=false>>=
options(SweaveHooks=list(fig=function() par(mar=c(2.5,2.5,0.1,0.1), mgp=c(1.5, 0.5, 0), cex=0.6)))
# source in all the R files
library(pandemic)
@

\begin{frame}
\frametitle{Swine Flu in Ontario}
\begin{itemize}
	\item Many Canadians holiday in Mexico during the winter and  spring

\begin{itemize}
	\item Swine flu was probably established in Ontario before the outbreak was detected.
\item Few control measures were adopted
\end{itemize}
\item Ontario's had recent public health crises
\begin{itemize}
	\item SARS, Walkerton water quality problems
	\item Ontario Agency for Health Protection and Promotion created
	\item Expert opinion abounds from mathematical modellers and infectious disease epidemiologists
	\item Detailed models are used for the biological process, with largely ad-hoc inference and treatment of uncertainty. 
\end{itemize}
\end{itemize}

\begin{frame}
\frametitle{The Data}
\begin{itemize}
	\item During the first 6 weeks of the epidemic, we can assume most cases were identified.

\begin{itemize}
	\item Ill individuals were encouraged to seek medical help
	\item Practitioners were encouraged to send samples to labs for testing
\end{itemize}
	\item After 6 weeks, testing was discouraged to ease the burden on public health labs.
	\item Roughly 1000 cases in the first 6 weeks? 
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{The Goal}
\begin{itemize}
	\item Provide good estimates (and inference) about hospitalization and fatality rates in the early stages of an epidemic
	\item To cope with sparse data, allow for expert opinion.
	\item Allow for variations in rates by age
	\item and account for individual's previous medical conditions?
\item Predict the number of hospital beds required for a given number of infections.
\end{itemize}
\end{frame}

@@

\begin{block}{Mild}
\begin{center}
 Infection $\stackrel{f_I(t)}{\Longrightarrow}$ Onset $\stackrel{f_{OM}(t)}{\Longrightarrow}$ \\ Medical  $\stackrel{f_{MR}(t)}{\Longrightarrow}$ Recovery
\end{center}

@@

\begin{frame}
\frametitle{The Data (again)}
\begin{itemize}
	\item Everyone has a date of a medical visit
	\item Some have date of onset of symptoms
	\item Most infections are mild and no additional data results
	\item Some end up in hospital, resulting in either recovery or death
	\item Longitudinal data on each case can be assembled\ldots in theory
\end{itemize}

\begin{frame}
\frametitle{Distributions}
\begin{itemize}
	\item $f(t;\mu,\nu,\delta)$ is the distribution for time to an event.  
	\item \ldots a rounded, zero-inflated Weibull
\begin{align*}
Z\sim&\text{Bern}(\delta)\\
[Y|Z=1] =& 0 \text{ w.p.} 1\\
pr(T=t|Z=0) =&  \text{pr}(\max(t-0.5,0) < y < t+0.5)\\
y \sim & \text{Weibull}(\mu/\Gamma(1 + 1/\nu), \nu)
\end{align*}
\item $\delta$ is the probability of transitioning immediately for an administrative (not biological) reason, 
\begin{itemize}
	\item i.e.\ arriving dead at the hospital.
\end{itemize}
\item Barring administrative zeros, $\mu = E(T|Z=0)$ is the average time to an event
\begin{itemize}
	\item $\mu_I$ is average time to infection
\end{itemize}
\item $\nu$ is a Weibull shape parameter
\begin{itemize}
	\item $\nu=1$ is constant hazards
	\item $\nu > 1$ results in more regular event times.
\end{itemize}

@@

\begin{frame}
\frametitle{Prior graphs}
<<priorGraph, fig=false,results=tex>>=
priors= readPrior("bayesPandemicPriors.txt")
plotPrior(priors$InfOns, file="InfOns", tex=TRUE)
#$
@

@@

\begin{frame}
\frametitle{To do}
\begin{itemize}
	\item simulate hospital loads
	\item model the infection process
	\item spatial infection process
	\item covariates

\end{itemize}
\end{frame}
